\NAME: A framework for \\ Efficient Intrusion Detection\\ in the Internet of Things

### 2014-11-21 paper.tex
fixed more typos and bad sentences

In classic networks, firewalls focus on the outer perimeter of the network,
filtering unwanted packets and protecting the entire internal network. But
attacks on flaws in services can pass unnoticed. This is where intrusion
detection (ID) comes into play: an intrusion detection systems (IDS) monitors
all traffic that passes through the firewall, looking for patterns of malicious
activity, and optionally, after detecting such pattern, alerting the firewall,
allowing it to take corrective actions \cite{denning1987intrusion}.

@@

\begin{figure}[ht]
  \centering
  \scalebox{.85}{
\begin{tikzpicture}[
  sibling distance=20mm,node distance=40mm
  ]
  \node (detection)             [align=center] {Detecting\\Intrusions}
    child {node (anomaly)       [align=center] {Anomaly}}
    child {node (signature)     [align=center] {Signature\\Misuse}}
    child {node (specification) [align=center] {Specification}};
  \node (decision)              [align=center,right=of detection] {Decision\\Making}
    child {node (broadcasting)  [align=center] {Broadcasting}}
    child {node (interactive)   [align=center] {Interactive}};
\end{tikzpicture}
}
  \caption{Classification of ID Algorithms}
  \label{fig:classification}
\end{figure}

@@

\begin{figure}[ht]
\captionof{lstlisting}{ID Algorithm Pattern\label{alg:id-algo-pattern}}
\begin{algorithmic}[1]
  \Require{devices: global storage for device information}
  \Function{process\_message}{$msg$}
    \ForEach{$byte \in msg$} \label{alg:id-algo-pattern-loop1}
     \State \dots \Comment{analyze byte-sequences}
    \EndFor
    \Let{$devices_x$}{value}  \Comment{optionally update device info}
    \State \Call{send}{$device_y$, ``info''} \Comment{optionally exchange info} \label{alg:id-algo-pattern-send1}
  \EndFunction
  \State
  \Function{do\_housekeeping}{}
    \ForEach{$device \in devices$} \label{alg:id-algo-pattern-loop2} \label{alg:id-algo-pattern-common-data}
      \If{$device > \dots$} \Comment{validate recorded value}
        \State \dots \Comment{take actions}
        \State \Call{send}{$device$, ``info''} \Comment{communicate} \label{alg:id-algo-pattern-send2}
      \EndIf
    \EndFor
  \EndFunction
  \State
\end{algorithmic}
\end{figure}

@@

Figure \ref{fig:positioning} shows a classification of possible development
strategies for projects involving wireless devices.

@@

Three properties can classify development strategies \cite{sadilek2007energy}:
flexibility, update cost and development effort.

@@

To convert the DSL-based ID algorithms into production-ready code, we apply
code generation techniques. Although different approaches here are applicable,
like template-based text expansion, XSLT transformations or CodeDom
\cite{dollard2004code}, we believe that the functional transformations require
a richer model.

@@

Figure \ref{fig:meta-model} shows the central part of the semantic model as
implemented for \NAME.

@@

The concept driving this model is the structure starting at the \emph{Model} up
to the \emph{Execution Strategy}. This structure encompasses the
\emph{Function}s and underlying \emph{Statement}s and allows for interpretation
and fusion of the algorithms.

@@

Thanks to the functional description of events and the effects triggered by
them, the code generator is able to identify the same actions across different
algorithms. This way it is possible to extract the parsing of incoming
messages, group actions on the same sets of data or scheduled function
executions.

FCF typically looks for patterns in the semantic model and performs model
transformations to change these patterns to a more optimized organization. It
first performs semantic transformations on the semantic model itself and ends
then transforms the semantic model into a syntactic model, encoding the
semantics into technical syntax constructs.

@@

Other common functionality also isn't generated as-is. The execution of
functions based on a schedule or the implementation of higher level conceptual
functions, such as \ttt{now} and \ttt{sha1}, are calls into the standard
software library that supports the code generator.

@@

To evaluate the prototype we chose to use a system based on the Atmel
ATMEGA1284p micro-controller and the Digi XBee S2 ZigBee module. A basic
library wrapped the technical calls to the hardware components in more easy to
use functions.

Relying on basic hardware and software allows us to validate that the generator
is capable of generating code even for the most basic environment available.
Any step up, both in hardware or software, would offer better and higher
abstractions that would make it easier to generate code for that situation. It
shows that the requirements of the generator towards the platform are minimal,
don't rely on advanced frameworks or operating systems and therefore are
applicable in any environment, targeting any platform.

@@

These results show that adding algorithms manually to the base-application has
a cumulative effect. The impact of both algorithms is simply the sum of the
impact of each algorithm by itself. In case of the execution time of one cycle
of the event loop, the total is even a bit higher.

@@

DSLs offer a solution here, but other paradigms are possible. A mixed framework
consisting of a virtual machine (VM) and native code extension
\cite{sadilek2007energy}, combined with energy-aware compilation, seems a
promising approach. By offloading the optimization to a VM and only compiling
parts and not the whole, the solution allows updating at a granular scale, thus
reducing the update cost, while maintaining flexibility and lowering the
development effort. As explained in \ref{positioning}, \NAME focuses on the
reduction of the development effort and update cost, while preserving maximal
flexibility.

>>>>

In classic networks, firewalls focus on the outer perimeter of the network,
filtering unwanted packets and protecting the entire internal network. But
attacks on flaws in services can pass unnoticed. This is where intrusion
detection (ID) comes into play: an intrusion detection system (IDS) monitors
all traffic that passes through the firewall, looking for patterns of malicious
activity, and optionally, after detecting such pattern, alerting the firewall,
allowing it to take corrective actions \cite{denning1987intrusion}.

@@

\begin{figure}[ht]
  \centering
  \scalebox{.85}{
\begin{tikzpicture}[
  sibling distance=20mm,node distance=40mm
  ]
  \node (detection)             [align=center] {Detecting\\Intrusions}
    child {node (anomaly)       [align=center] {Anomaly}}
    child {node (signature)     [align=center] {Signature\\Misuse}}
    child {node (specification) [align=center] {Specification}};
  \node (decision)              [align=center,right=of detection] {Decision\\Making}
    child {node (broadcasting)  [align=center] {Broadcast}}
    child {node (interactive)   [align=center] {Interactive}};
\end{tikzpicture}
}
  \caption{Classification of ID Algorithms}
  \label{fig:classification}
\end{figure}

@@

\begin{figure}[ht]
\captionof{lstlisting}{ID Algorithm Pattern\label{alg:id-algo-pattern}}
\begin{algorithmic}[1]
  \Require{devices: global storage for device information}
  \Function{process\_message}{$msg$}
    \ForEach{$byte \in msg$} \label{alg:id-algo-pattern-loop1}
     \State \dots \Comment{analyze byte-sequences}
    \EndFor
    \Let{$devices_x$}{value}  \Comment{optionally update device info}
    \State \Call{send}{$devices_y$, ``info''} \Comment{optionally exchange info} \label{alg:id-algo-pattern-send1}
  \EndFunction
  \State
  \Function{do\_housekeeping}{}
    \ForEach{$device \in devices$} \label{alg:id-algo-pattern-loop2} \label{alg:id-algo-pattern-common-data}
      \If{$device > \dots$} \Comment{validate recorded value}
        \State \dots \Comment{take actions}
        \State \Call{send}{$device$, ``info''} \Comment{communicate} \label{alg:id-algo-pattern-send2}
      \EndIf
    \EndFor
  \EndFunction
  \State
\end{algorithmic}
\end{figure}

@@

@@

Figure \ref{fig:positioning} shows a classification of possible development
strategies for projects involving wireless devices. Three properties can
classify development strategies \cite{sadilek2007energy}: flexibility, update
cost and development effort.

@@

To convert the DSL-based ID algorithms into production-ready code, \NAME
applies code generation techniques. Although different approaches here are
applicable, like template-based text expansion, XSLT transformations or CodeDom
\cite{dollard2004code}, we believe that the functional transformations require
a richer model.

@@

@@

Figure \ref{fig:meta-model} shows the central part of the semantic model as
implemented for \NAME. The concept driving this model is the structure starting
at the \emph{Model} up to the \emph{Execution Strategy}. This structure
encompasses the \emph{Function}s and underlying \emph{Statement}s and allows
for interpretation and fusion of the algorithms.

@@

Thanks to the functional description of events and the effects triggered by
them, the code generator is able to identify the same actions across different
algorithms. This way it is possible to extract the parsing of incoming messages
and group actions on the same sets of data or scheduled function executions.

FCF typically looks for patterns in the semantic model and performs model
transformations to change these patterns to a more optimized organization. It
first performs semantic transformations on the semantic model itself. A second
phase transforms the semantic model into a syntactic model, encoding the
semantics into technical syntax constructs.

@@

Other common functionality also isn't generated as is. The execution of
functions based on a schedule or the implementation of higher level conceptual
functions, such as \ttt{now} and \ttt{sha1}, are calls into the standard
software library that supports the code generator.

@@

To evaluate the prototype we used a system based on the Atmel ATMEGA1284p
micro-controller and the Digi XBee S2 ZigBee module. A basic library wrapped
the technical calls to the hardware components in more easy to use functions.

Relying on basic hardware and software allows us to validate that the generator
is capable of generating code even for the most basic environment. Any step up,
both in hardware or software, would offer better and higher abstractions that
would make it easier to generate code for that situation. It shows that the
requirements of the generator towards the platform are minimal, don't rely on
advanced frameworks or operating systems and therefore are applicable in any
environment, targeting any platform.

@@

These results show that adding algorithms manually to the base-application has
a cumulative effect. The impact of both algorithms is simply the sum of the
impact of each algorithm by itself. The execution time of one cycle of the
event loop with both algorithms enabled is even higher than simply the sum.

@@

DSLs offer a solution here, but other paradigms are possible. A mixed framework
consisting of a virtual machine (VM) and native code extension
\cite{sadilek2007energy}, combined with energy-aware compilation, seems a
promising approach. By offloading the optimization to a VM and only compiling
parts and not the whole, the solution allows updating at a granular scale, thus
reducing the update cost, while maintaining flexibility and lowering the
development effort. As explained in section \ref{positioning}, \NAME focuses on
the reduction of the development effort and update cost, while preserving
maximal flexibility.
